Hippocampal replay of experience at real-world speeds

### 2020-10-05 Article.tex

%TC:break Abstract
%the command above serves to have a word count for the abstract
\begin{abstract}
Representations of past and possible future experiences play a critical role in memory and decision-making processes. The hippocampus expresses these sorts of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at ~20x real-world speeds. Efforts to understand replay typically make multiple assumptions about which events to examine and what sorts of representations constitute replay, however. We therefore lack a clear understanding of both the prevalence and the range of representational dynamics associated with replay. Here we develop a state space model that uses a combination of movement dynamics of different speeds to capture the spatial content and time evolution of replay during SWRs. We find that the large majority of replay events contain spatially coherent, interpretable content. Furthermore, most events progress at real-world, rather than accelerated, movement speeds, consistent with actual experiences.

@@

Hippocampal "replay" events are well suited to that role \cite{CarrHippocampalreplayawake2011}. As animals move through space, neurons in the hippocampus preferentially fire at specific locations in an environment, and thus sets of cells fire in sequence as the animal moves through a series of locations. When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event. A subset of SWRs contain sequential firing similar to that seen during a previous experience, and are thus thought to "replay" these previous experiences. Importantly, previous work has reported that these sequential firing events proceed at an average speed of ~10 meters per second, about 20x faster than the animal's usual movement speed \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.

While the existence of these sequential events is well established, the current consensus is that only a minority (perhaps ~10-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay. Of the remaining events, some contain spiking of neurons associate with single locations where animals are immobile (Cite Yu et. al.), but another report suggested that events that replay a single location are only seen in young animals (Cite Stella). Thus, while it would seem useful to replay memories associated with single locations and perhaps even experiences occurring at real-world speeds, the prevalence of that sort of replay in awake animals remains unclear.

Our uncertainty stems in part from the dominant approaches used to identify the content of replay events. These approaches typically involve multiple steps and multiple assumptions about the nature of replay, which are mostly commonly characterized using the standard "Bayesian" decoder. First, an encoding model is constructed based on spiking activity during movement, most often using only spikes that have been clustered as single units (putative single neurons). Then, a subset of SWRs or events with high activity levels are selected based on a threshold for event size chosen by the experimenter \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009, StellaHippocampalReactivationRandom2019}. A decoding algorithm is then applied to the  spikes within these events, yielding a set of probability distributions. Current approaches use either overlapping or non-overlapping time bins whose size is also chosen by the experimenter. Finally, the most commonly used approach to detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant, non-zero speed \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009}. A statistical test is then used to determine whether the fit is better than the fit to shuffled versions of the data.

>>>>

%TC:break Abstract
%the command above serves to have a word count for the abstract
\begin{abstract}
Representations of past and possible future experiences play a critical role in memory and decision-making processes. The hippocampus expresses these types of representations during sharp-wave ripple (SWR) events, and previous work identified a minority of SWRs that contain “replay” of spatial trajectories at ~20x real-world speeds. Efforts to understand replay typically make multiple assumptions about which events to examine and what sorts of representations constitute replay. We therefore lack a clear understanding of both the prevalence and the range of representational dynamics associated with replay. Here we develop a state space model that uses a combination of movement dynamics of different speeds to capture the spatial content and time evolution of replay during SWRs. We find that the large majority of replay events contain spatially coherent, interpretable content. Furthermore, most events progress at real-world, rather than accelerated, movement speeds, consistent with actual experiences.

@@

Hippocampal "replay" events are well suited to that role \cite{CarrHippocampalreplayawake2011}. As animals move through space, neurons in the hippocampus preferentially fire at specific locations in an environment, and thus sets of cells fire in sequence as the animal moves through a series of locations. When the animal is asleep or immobile, hippocampal cells can be reactivated during a "sharp-wave ripple" (SWR) event. A subset of SWRs contain sequential firing similar to that seen during a previous experience, and are thus thought to "replay" these previous experiences. Importantly, previous work has reported that these sequential firing events proceed at an average speed of \texttildelow10 meters per second, about 20x faster than the animal's usual movement speed \cite{NadasdyReplayTimeCompression1999, LeeMemorySequentialExperience2002, DavidsonHippocampalReplayExtended2009, KarlssonAwakereplayremote2009}.

While the existence of these sequential events is well established, the current consensus is that only a minority (\texttildelow10-45\%) of hippocampal SWRs contain statistically identifiable, sequential replay. Of the remaining events, some contain spiking of neurons associated with single locations where animals are immobile \cite{JaiDistincthippocampalcorticalmemory2017}, but another report suggested that events that replay a single location are only seen in young animals \cite{StellaHippocampalReactivationRandom2019}. Thus, while it would seem useful to replay memories associated with single locations and perhaps even experiences occurring at real-world speeds, the prevalence of that type of replay in awake animals remains unclear.

Our uncertainty stems in part from the dominant approaches used to identify the content of replay events. These approaches typically involve multiple steps and multiple assumptions about the nature of replay, which are mostly commonly characterized using the standard "Bayesian" decoder. First, an encoding model is constructed based on spiking activity during movement, most often using only spikes that have been clustered as single units (putative single neurons). Then, a subset of SWRs or events with high activity levels are selected based on a threshold for event size chosen by the experimenter \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009, StellaHippocampalReactivationRandom2019}. A decoding algorithm is then applied to the spikes within these events, yielding a set of position probability distributions for each time bin. Current approaches use either overlapping or non-overlapping time bins whose size is also chosen by the experimenter. Finally, the most commonly used approach to detecting sequential replay involves fitting a line to the resulting set of probability distributions, which relies on the assumption that the representations progress at a constant, non-zero speed \cite{FosterReversereplaybehavioural2006, DibaForwardreversehippocampal2007, KarlssonAwakereplayremote2009}. A statistical test is then used to determine whether the fit is better than the fit to shuffled versions of the data.
